---
title: "Get iSDA Data"
output: html_notebook
author: pme

---

Get [iSDA](https://www.isda-africa.com/isdasoil/) data for the AOI, and draw some quick maps.
Following Tomislav Hengl's [directions](https://gitlab.com/openlandmap/africa-soil-and-agronomy-data-cube).  Layers can be found on [zenodo](https://zenodo.org/search?page=1&size=50&q=iSDAsoil).

# Setup

## Variables and local data
Note to specify overwriting
```{r}
set_crs = '+init=epsg:4326'  # WGS84
set_projection = '+init=epsg:32736'  # utm 36S. For setting a buffer around the points.

# locations
in_dir = 'Data'
dir_local = file.path(in_dir, 'Spatial')
in_pt = 'Empirical Points.gpkg'  # cleaned up empirical data. Combined 2016 PANEL with later bulk density. 

# mwi bounds. Download from the repository of your choice - ex. Malawi geospatial data portal. 
in_bounds = "Malawi gadm36_MWI.gpkg"

# Land suitability - provided in file
in_suit = "LI_ETAL_2017_MW_AGRI_LAND_SUIT.tif"

# DEM. Download from the source of your choice, example Google Earth.
in_dem = "SRTM90_V4_MWI.tif"  # from Google Earth Engine via ree

# output
out_ras = 'ISDA_AR_MWI.tif'
out_pt = 'Master Points.gpkg'
out_dem = 'srtm90_MWI.tif'
out_suit = 'ag_suitability_wgs84.tif'
overwrite = FALSE
```

## iSDA
Set up iSDA downloads, including unit conversions. Need to specify:

1. Base URL preceded by vsicurl, which package `terra` recognizes
2. Layers to access. The part of the name preceding '_m.....' can be found on zenodo.
3. Transformations for each layer. Follow the format of the `switch()` function within the 
local function `transform_isda()`. Again, see the zenodo site for each layer. 
4. Units for each layer. 
```{r}
url_isda = '/vsicurl/https://s3.eu-central-1.wasabisys.com/africa-soil/layers30m/'


in_isda = c(TC = 'sol_log.c_tot',      # convert by expm1(x/10)
            SOC = 'sol_log.oc',        # convert by expm1(x/10)
            BD = 'sol_db_od',          # requires multiplying by 10 to get kg m^-3.
            PH = 'sol_ph_h2o',         # requires dividing by 10 to get pH units
            SAND = 'sol_sand_tot_psa',
            SILT = 'sol_silt_tot_psa',
            CLAY = 'sol_clay_tot_psa')
in_isda_tag = paste0('_m_30m_0..20cm_2001..2017_africa_epsg4326_v0.1.tif')


in_isda_units = c(
  TC = 'g/kg',
  SOC = 'g/kg',
  BD = 'g/cm3',
  PH = 'pH',
  SAND = 'g/g',
  SILT = 'g/g',
  CLAY = 'g/g')
```

Some iSDA layers need transforming to SI-adjacent units. 
```{r}
# Identify transformations here.
transform_isda = function(layer, layer_name, units) {
  if (is.null(attr(layer, 'units', exact=TRUE))) {
    out = 
      switch(layer_name,
             'TC' = expm1(layer/10),
             'SOC' = expm1(layer/10),
             'BD' = layer/100,
             'PH' = layer/10,
             'SAND' = layer/100,
             'SILT' = layer/100,
             'CLAY' = layer/100)
    
    attr(out, 'units') = units[layer_name]
  } else {
    out = layer
  }
  
  return(out)
}

```

## Libraries
```{r}
libs = c('sf',        
         'terra', 
         'rgdal',
         'magrittr',
         'car',
         'ggplot2',
         'here')  
```

# Begin
## Load libraries
```{r message=FALSE}
for (i in libs) {
  if (!require(i, character.only=TRUE)) {
    renv::install(i)
    require(i, character.only=TRUE)
  }}

set_crs %<>% CRS
set_projection %<>% CRS
```

## Set up AOI
Include a 20 km buffer around points.
```{r}
pts = here(in_dir, in_pt) %>% 
  st_read
  # read.csv(header=TRUE) %>% 
  # st_as_sf(coords=in_pt_coords,
  #          crs=set_crs) 

mwi = here(dir_local, in_bounds) %>% 
  st_read(layer='gadm36_MWI_1') %>% 
  .['NAME_1']

ggplot(mwi) +
  geom_sf() +
  geom_sf(data=pts,
          col='black',
          size=0.2) +
  labs(title='AOI') +
  theme_minimal()
```

# Get iSDA
crop to AOI and transform. This is slow, so broken down pretty carefully.

First point terra to the layers on Zenodo
```{r}
isda_names = names(in_isda) %>% 
  set_names(., .)

ISDA = lapply(isda_names, function(x) {
  # get the layer
  paste('Working on', x) %>% 
    message
  layer = paste0(url_isda, in_isda[x], in_isda_tag) %>% 
    rast %>% 
    crop(mwi) %>% 
    mask(mwi) %>% 
    transform_isda(x, in_isda_units)
  return(layer)
})
```

# Combine into a raster stack and plot
```{r}
out = ISDA[[1]]
for (i in 2:length(ISDA)) {
  terra::add(out) = ISDA[[i]]
}

names(out) = paste(
  isda_names,
  sapply(ISDA, attr, 'units'),
  sep='_')

# if this fails, then run before repeating this cell:
# ISDA[[1]] = ISDA[[1]][[1]]
```


```{r}
plot(out)
```


```{r}
here(dir_local, 'ISDA_AR_MWI.png') %>% 
  png(height=11, width=8.5, units='in', res=300)
plot(out)
dev.off()

```


## Export as a geotiff
```{r}
# rename layers as ISDA_<property>
names(out) %<>% strsplit('_') %>% 
  sapply('[', 1) %>% 
  paste0('ISDA_', .)
here(dir_local, out_ras) %>% 
  writeRaster(out, ., overwrite=overwrite)
```

# DEM
```{r}
dem = here(in_dir, dir_spatial, in_dem) %>% 
  rast %>% 
  crs_check(set_datum) %>% 
  crop(mwi) %>% 
  mask(mwi)
plot(dem)
```
## Export
```{r}
here(dir_local, out_dem) %>% 
  writeRaster(dem, ., overwrite=overwrite)
```


# Land suitability
Reproject, resample, and export
```{r}
suit = here(dir_local, in_suit) %>% 
  rast
suit_mask = suit < 5
suit_mask[!suit_mask] = NA
terra::add(suit) = suit_mask
names(suit) = c('suitability', 'masked')
suit =
  terra::project(suit, out, method='near', align=TRUE, mask=TRUE) %>% 
  crop(out)
plot(suit)
```

## Export
```{r}
here(dir_local, out_suit) %>% 
  writeRaster(suit, ., overwrite=overwrite)
```


# Extract values at points
```{r}
pts_vect = vect(pts)
for (x in isda_names) {
  plt_name = paste0('ISDA_', x)
  pts[, plt_name] = terra::extract(ISDA[[x]], pts_vect)[, 2]
}
pts[, 'ELEVATION'] = terra::extract(dem, pts_vect)[, 2]
pts[, 'SUITABILITY'] = terra::extract(suit[[1]], pts_vect)[, 2]

# convert bulk density to g/cm3
pts[, 'BULK_DENSITY'] %<>% divide_by(1000)

# convert texture to g/g
pts[, 'SAND'] %<>% divide_by(100)
pts[, 'SILT'] %<>% divide_by(100)
pts[, 'CLAY'] %<>% divide_by(100)
```

## Plot
```{r}
is_isda = grepl('ISDA_', names(pts)) %>% 
  names(pts)[.] %>% 
  c('ELEVATION', 'SUITABILITY')
plot(pts[is_isda])
```


```{r fig.height=8, fig.width=8, warning=FALSE, message=FALSE}
as.data.frame(pts) %>% 
  .[sample(seq_len(nrow(pts)), 300), is_isda] %>% 
  scatterplotMatrix
```


## Export
```{r}
here(in_dir, out_pt) %>% 
  st_write(pts, ., delete_layer=overwrite)
```

```{r}

```

