---
title: "Fig 1 - Study Site"
knit: (function(inputFile, encoding) {
  rmarkdown::render(inputFile, encoding=encoding, output_dir=here::here('Results'))})
output:
  html_document:
    toc: true
    toc_depth: 5
    toc_float: true
    df_print: paged
    code_folding: hide
author: PME
---

Map of malawi, sample locations, and values of critical soil/edaphic properties in relation to other agricultural lands.

# Load

## Paths, Variables, Libraries

Variables
```{r}
set_datum = 'epsg:4326'  # WGS84
set_projection = 'epsg:32736'  # utm 36S.
# output
overwrite = FALSE
# overwrite = TRUE
```

Paths
```{r}
# locations
in_dir = 'Data'
in_pt = 'Master Points.gpkg'
# in_pt = 'AR_MWI_ISDA_points.gpkg'

dir_spatial = 'Spatial' # in in_dir
in_isda = 'ISDA_AR_MWI.tif'
# in_suit = 'LI_ETAL_2017_MW_AGRI_LAND_SUIT.tif'  
in_suit = "ag_suitability_wgs84.tif" # cropland suitability
in_mwi = "Malawi gadm36_MWI.gpkg"
in_dem = "srtm90_MWI.tif"  # from Google Earth Engine via ree

# in_isda = c(
#   'BD',
#   'CLAY',
#   'SAND',
#   'SILT',
#   'PH',
#   'SOC')
# in_isda = paste0("ISDA_", in_isda)
# names(in_isda) = in_isda

out_dir = 'Results'
out_fig = 'Figure 1 - Study Area and Properties.jpg'

```

Libraries
```{r}
libs = c(
  # Coding helpers
  'here', 
  'tidyr',
  'magrittr',
  
  # spatial stuff
  'sf',
  'terra',
  'stars',
  
  # Other modeling
  'glmnet',
  'quantreg',
  
  # plotting
  'ggplot2',
  'rasterVis',
  'gridExtra'
)
```

```{r message=FALSE}
for (i in libs) {
  if (!require(i, character.only=TRUE)) {
    renv::install(i)
    require(i, character.only=TRUE)
}}

resource = function() {
  ll = here('Scripts', 'V3 Functions') %>% 
  list.files(full.names=TRUE)
  for (i in ll) {
    source(i)
  }
}
resource()
```

Versions
```{r}
sapply(libs, function(x) as.character(packageVersion(x))) %>% 
  as.matrix %>% 
  set_colnames(R.version.string)
```

## Load

### Point Data
```{r cache=TRUE}
pt = here(in_dir, in_pt) %>% 
  st_read
```

### iSDA
load and convert units
```{r fig.height=4, fig.width=8, warning=FALSE}
isda = #paste0(in_isda, '_MWI.tif') %>% 
  here(in_dir, dir_spatial, in_isda) %>% 
  rast

summary(isda)
```
Looks good. 

### Malawi EPA bounds
```{r}
mwi = here(in_dir, dir_spatial, in_mwi) %>% 
  st_read(layer='gadm36_MWI_1')
```

### Elevation
```{r}
dem = here(in_dir, dir_spatial, in_dem) %>% 
  rast
```

### Suitability 
```{r}
suit = here(in_dir, dir_spatial, in_suit) %>% 
  rast
```


### Check CRS and combine

Add the DEM layer to iSDA for convenience - but keep effective resolution
```{r}
isda$ELEVATION = resample(dem, isda, method='near')
```

Project suitability and make a mask.
```{r}
suit_mask = suit[[2]]
plot(suit_mask, col=c('red', 'white'))
suit = suit[[1]]
```

### A few more functions
```{r}
gg_mapthemer = function(x, xlab='Long (E)', ylab='Lat (S)',
                        coord_equal=FALSE) {
  out = x +
    labs(x=xlab,
         y=ylab) +
    theme_minimal() +
    theme(axis.text.x=element_text(angle=90, 
                                   hjust=1, 
                                   vjust=0.5))
  if (coord_equal) {
    out = out + coord_equal()
  }
  return(out)
}
```

# Map of study area
```{r}
mwi$LABEL = mwi$NAME_1 %>% 
  sapply(function(x) ifelse(x %in% c('Dedza', 'Machinga', 'Ntcheu'), x, NA))

pltdf = aggregate(isda, fact=10)
# is_water = pltdf$ISDA_BD == 0
# is_water[!is_water] = NA
# is_water %<>% 
#   st_as_stars %>% 
#   st_as_sf(as.points=FALSE, merge=TRUE)

focus_epa = mwi[!is.na(mwi$LABEL), ]

suit_agg = aggregate(suit, 10, fun='modal') %>% 
  resample(pltdf, method='near') %>% 
  mask(pltdf[[1]])
```


```{r fig.height=8, fig.width=4}
study_area = 
  ggplot(mwi) +
  # coord_sf(crs=st_crs(set_projection)) +
  # political boundaries as background
  scale_x_continuous(breaks=seq(33, 36, by=1)) +
  scale_fill_distiller(palette='Greens', na.value=NA,
                       limits=c(0.9, 5.1),
                       direction=-1) +
  scale_color_manual(values=c('tomato1', 'tomato3', 'tomato4')) +
  
  # stock level
  # new_scale('fill') +
  # water base layer
  geom_sf(fill='skyblue1', 
          color='skyblue1') +
 
  # suitability
  geom_stars(data=st_as_stars(suit_agg),
             alpha=0.8) +
  
  # political boundaries
  geom_sf(color='#666666',
          alpha=0,
          size=0.2) +
  
  # focus EPAs
  geom_sf(data=focus_epa,
          aes(color=LABEL),
          fill=NA,
          # alpha=0.2,
          size=0.6) +
  geom_sf(data=pt,
          color='black',
          size=0.2) +
  labs(x=NULL,
       y=NULL,
       fill='Land Suitability\nClass',
       color='District',
       tag='a)') +
  theme_minimal() +
  theme(legend.position='bottom',
        legend.direction='vertical',
        axis.text=element_text(size=8),
        legend.title=element_text(size=8),
        legend.text=element_text(size=8),
        title=element_text(size=9))
study_area
```

# Representativeness of points.

## Sample from across Malawi
```{r}
spl = mask(isda, suit_mask) %>% 
  spatSample(5000, as.points=TRUE, na.rm=TRUE)
spl$SAMPLEID = paste0('S', seq_len(nrow(spl)))
```

```{r}
plot(isda$ELEVATION)
points(spl, pch='.')
lines(vect(mwi))
```

## Plot
```{r}
pltdf = data.frame(pt) %>% 
  .[, names(spl)] %>% 
  pivot_longer(cols=where('is.numeric'))
pltdf$SOURCE = 'Visited'

spl %<>% data.frame %>% 
  pivot_longer(cols=where('is.numeric'))
spl$SOURCE = 'Nation-wide'

pltdf %<>% rbind(spl)

prettynames = c(
  SAND = expression(paste('Sand Content [g', ~g^-1, ']', sep='')),
  SILT = expression(paste('Silt Content [g', ~g^-1, ']', sep='')),
  CLAY = expression(paste('Clay Content [g', ~g^-1, ']', sep='')),
  BD = expression(paste('Bulk Density [g', ~cm^-3, ']', sep='')),
  PH = expression(paste('pH')),
  SOC = expression(paste('Organic Carbon [g', ~kg^-1, ']', sep='')),
  ELEVATION = expression(paste('Elevation [m]'))
)
pltdf$name %<>% 
  gsub('ISDA_', '', .) %>% 
  prettynames[.] %>% 
  factor(., levels=prettynames)
drop = is.na(pltdf$name)
pltdf %<>% .[!drop, ]
```

```{r fig.height=6, fig.width=4}
scale_setter = data.frame(SAMPLEID=NA,
                          name='SOC',
                          value=0,
                          SOURCE=NA)
scale_setter$name %<>%
  prettynames[.] %>% 
  factor(levels=prettynames)

property_plot = 
  ggplot(pltdf,
       aes(x=value,
           color=SOURCE)) +
  scale_color_manual(values=c("#BBBBBB", "#666666")) +
  facet_wrap(~name, 
             scales='free', 
             ncol=2, 
             strip.position='bottom',
             labeller='label_parsed') +
  geom_blank(data=scale_setter) +
  geom_density(adjust=1.7) +
  labs(x=NULL,
       y='Density',
       color=NULL,
       tag='b)') +
  geom_hline(yintercept=0,
             color='lightgray',
             size=0.2) +
  theme_minimal() +
  theme(legend.position=c(0.75, 0.05),
        axis.text=element_text(size=8),
        title=element_text(size=9),
        legend.text=element_text(size=8),
        strip.text=element_text(size=8, vjust=1),
        strip.placement='outside',
        panel.grid=element_blank(),
        axis.ticks=element_line(color='lightgray', 
                                size=0.2))
property_plot
```

```{r fig.height=6, fig.width=6.5}
grid.arrange(grobs=list(study_area, property_plot),
             widths=c(2,3), 
             nrow=1)
```
# Export
```{r}
here(out_dir, out_fig) %>% 
  jpeg(width=6.5, height=6.5, units='in', res=300)
grid.arrange(grobs=list(study_area, property_plot),
             widths=c(2,3), 
             nrow=1)
dev.off()
```

