---
title: "Fig 4 - Updated maps of stocks and gaps"
knit: (function(inputFile, encoding) {
  rmarkdown::render(inputFile, encoding=encoding, output_dir=here::here('Results'))})
output:
  html_document:
    toc: true
    toc_depth: 5
    toc_float: true
    df_print: paged
    code_folding: hide
author: PME
---

Run this after Part 3. 

In this notebook, we will produce updated maps of carbon stocks, storage potentials,
and gaps. 

The prediction model used will be the top-down stationary model, which performed
equivalently to the spatial model at unobserved regions and performed better than 
the bottom-up model.

Goals are:
1. Predict carbon stocks across the region through updating iSDA
2. Predict storage potential based on soil properties
3. Calculate carbon gaps for agricultural land.


# Load

## Paths, Variables, Libraries

Variables
```{r}
set_datum = 'epsg:4326'  # WGS84
set_projection = 'epsg:32736'  # utm 36S.
# output
overwrite = FALSE
# overwrite = TRUE

plot_diagnostics = TRUE
```

Paths
```{r}
# locations
in_dir = 'Data'
in_pt = 'Master Points.gpkg'

dir_spatial = 'Spatial' # in in_dir
in_isda = 'ISDA_AR_MWI.tif'
in_suit = 'LI_ETAL_2017_MW_AGRI_LAND_SUIT.tif'  # cropland suitability
in_mwi = "Malawi gadm36_MWI.gpkg"
in_dem = "srtm90_MWI.tif" 

out_dir = 'Results'

# in_isda = c(
#   'BD',
#   'CLAY',
#   # 'SAND',
#   'SILT',
#   # 'PH',
#   'SOC')
# in_isda = paste0("ISDA_", in_isda)
# names(in_isda) = in_isda
# 
# in_isda = 'ISDA_AR_MWI.png'

```

Libraries
```{r}
libs = c(
  # Coding helpers
  'here', 
  # 'tidyr',
  'magrittr',
  
  # spatial stuff
  'sf',
  'terra',
  'stars',
  
  # Other modeling
  'glmnet',
  'quantreg',
  
  # plotting
  'ggplot2',
  'rasterVis',
  'gridExtra'
)
```

```{r message=FALSE}
for (i in libs) {
  if (!require(i, character.only=TRUE)) {
    renv::install(i)
    require(i, character.only=TRUE)
}}

resource = function() {
  ll = here('Scripts', 'V3 Functions') %>% 
  list.files(full.names=TRUE)
  for (i in ll) {
    source(i)
  }
}
resource()
```

Versions
```{r}
sapply(libs, function(x) as.character(packageVersion(x))) %>% 
  as.matrix %>% 
  set_colnames(R.version.string)
```

## Load
```{r}
load(here(in_dir, 'Fig 2 - Field Estimates.Rdata'))
load(here(in_dir, 'Part 3 - iSDA Corrections.Rdata'))
```

### Malawi EPA bounds
```{r}
mwi = here(in_dir, dir_spatial, in_mwi) %>% 
  st_read(layer='gadm36_MWI_1')
```

### iSDA
load and convert units
```{r fig.height=4, fig.width=8, warning=FALSE}
isda =  here(in_dir, dir_spatial, in_isda) %>% 
  rast
summary(isda)
```
Looks good. 

### Elevation
Add directly to iSDA data
```{r}
isda$ELEVATION = here(in_dir, dir_spatial, in_dem) %>% 
  rast %>% 
  resample(isda, method='near')
```

### Mask to land suitable for agriculture
Make a mask that aligns with stocks. Must be in WGS84 with 30m resolution. 
```{r}
suit_mask = here(in_dir, dir_spatial, in_suit) %>% 
  rast
suit_mask = suit_mask < 5
suit_mask[!suit_mask] = NA
suit_mask %<>% 
  project(isda, align=TRUE, method='near') %>% 
  crop(isda)

isda %<>% mask(suit_mask)
```
#### Export iSDA for the adjustment demo
```{r}
here(dir_spatial, "iSDA_combined_masked.tif") %>% 
  writeRaster(isda,
              .,
              overwrite=overwrite)
```



### Reproject boundaries
```{r}
mwi_utm = st_transform(mwi, set_projection)
```


### A few more functions
```{r}
gg_mapthemer = function(x, 
                        xlab=NULL, 
                        ylab=NULL,
                        coord_equal=FALSE) {
  out = x +
    labs(x=xlab,
         y=ylab) +
    theme_minimal() +
    theme(axis.text.x=element_text(angle=90, 
                                   hjust=1, 
                                   vjust=0.5))
  if (coord_equal) {
    out = out + coord_equal()
  }
  return(out)
}
```


```{r}
ggmap = function(raster, fill='value', label=NULL) {
  require(rasterVis)
  out = gplot(raster) +
    geom_tile(aes_string(fill=fill)) +
    scale_fill_continuous(na.value=NA) +
    coord_equal() +
    labs(fill=label)
  out = gg_mapthemer(out)
  return(out)
}
ggmap(isda$ISDA_SOC, label='Soil Organic\nCarbon [g/kg]')
```


# Estimate stocks from iSDA corrections
## Calculate current iSDA Stock Estimates

Excluding "inorganic" carbon.
```{r}
isda$ISDA_CSTOCK_NOW = calc_stock(isda$ISDA_SOC/1000, isda$ISDA_BD, 0.2)
```

## Adjust isda estimates. Place in a new object, "out_stocks". 
```{r}
pred = 'CSTOCK_NOW'
model = current_stocks[['top down stationary']]

predict_se = function(...) predict(..., type='response', se.fit=TRUE)$se.fit

out_stocks = predict(isda,
                     model,
                     type='response') %>% 
  set_names(pred)
se = predict(isda, 
             model,
             fun=predict_se) %>% 
  set_names(paste0(pred, '_se'))

terra::add(out_stocks) = se
```



```{r fig.height=5, fig.width=8}
grid.arrange(
  ggmap(out_stocks$CSTOCK_NOW, 
        label='Corrected\nCarbon Stock\n[Mg/ha]'),
  ggmap(out_stocks$CSTOCK_NOW_se, 
        label='Corrected\nCarbon Stock\nError [Mg/ha]'),
  nrow=1
)
```

## Calculate potentials
Layer called CSTOCK_80. Set floor of current stocks. 
```{r}
pred = 'CSTOCK_80'
model = potential_stocks[['top_stat']]

# predict
new_layers = predict(isda, 
                     model,
                     fun=rast_predict_se)

names(new_layers) = c(
  pred,
  paste0(pred, '_se')
)

# Set floor
tt = 
  max(new_layers[[pred]], 
      out_stocks[['CSTOCK_NOW']])


terra::add(out_stocks) = tt
terra::add(out_stocks) = new_layers[[2]]
```


```{r fig.height=5, fig.width=8}
grid.arrange(
  ggmap(out_stocks$CSTOCK_80, 
        label='Potential\nCarbon Stock\n[Mg/ha]'),
  ggmap(out_stocks$CSTOCK_80_se, 
        label='Potential\nCarbon Stock\nError [Mg/ha]'),
  nrow=1
)
```


## Calculate gaps
Layer called CGAP. Min is 0 because of floor for potential stocks.
```{r}
gap = out_stocks[['CSTOCK_80']] - out_stocks[['CSTOCK_NOW']]
names(gap) = 'CGAP'
terra::add(out_stocks) = gap


gap_se = sqrt(out_stocks$CSTOCK_80_se^2 + out_stocks$CSTOCK_NOW_se^2)
names(gap_se) = 'CGAP_se'
terra::add(out_stocks) = gap_se
```


```{r fig.height=5, fig.width=8}
grid.arrange(
  ggmap(out_stocks$CGAP, 
        label='Carbon Gap\n[Mg/ha]'),
  ggmap(out_stocks$CGAP_se, 
        label='Carbon Gap\nError [Mg/ha]'),
  nrow=1
)
```

## Correlation between stocks and gaps
```{r}
tt = spatSample(out_stocks, 1000, na.rm=TRUE) %>% 
  .[, c('CSTOCK_NOW', 'CGAP')] %>% 
  cor %>% 
  round(3)
```

## Export
```{r}
here(out_dir, 'corrected_SOC_stocks_mwi.tif') %>%
  writeRaster(out_stocks,
              .,
              overwrite=overwrite)
out_stocks = here(out_dir, 'corrected_SOC_stocks_mwi.tif') %>% 
  rast
```

# Map
## Total land area and cell size
For calculating totals.
```{r}
suit_area = cellSize(suit_mask, unit='ha') %>% 
  global(fun='sum', na.rm=TRUE)

ggmap(suit_mask) +
  labs(title=paste("Area:", round(suit_area/1e6, 1), 'M ha'))
```

Cell size for calculating to total 
```{r}
cellsize = cellSize(out_stocks$CSTOCK_NOW, 
                    mask=TRUE, 
                    unit='ha')
ggmap(cellsize, label='Cell Size [ha]')
```

# Figure 4 - Estimated Stocks
Various pieces:

- aggregated data
- new, nicer names
- subtitles of totals across the region with errors. 
- breaks and ranges for color scales - to ensure stocks and potentials are on
on the same scale.

```{r fig.height=8}
data = out_stocks
is_se = grepl('_se', names(data))

pltdf = data[[!is_se]] %>% 
  aggregate(fact=10) %>% 
  project(set_projection)

nn = names(pltdf)

prettynames = c(
  CSTOCK_NOW = 'Current Stock',# [Mg C/ha]',
  CSTOCK_80 = 'Potential Stock',
  CGAP = 'Current Gap',
  CSTOCK_NOW_se = 'Current Stock Error',
  CSTOCK_80_se = 'Potential Stock Error',
  CGAP_se = 'Current Gap Error'
)

# calculates totals with errors. Gap totals are with a floor of zero. 
generate_subtitle = function(data, x) {
  total = data[[x]] %>% 
    multiply_by(cellsize) %>% 
    global(fun='sum', na.rm=T) %>% 
    divide_by(1e6) %>% 
    round(1) 
  
  error = paste0(x, '_se') %>% 
    data[[.]] %>% 
    {.^2} %>% 
    multiply_by(cellsize) %>% 
    global(fun='mean', na.rm=TRUE) %>% 
    sqrt %>% 
    multiply_by(suit_area) %>% 
    divide_by(1e6) %>% 
    round(1)
  
  if (grepl('se', x)) {
    out = NULL
  } else {
    out = paste(total, '±', error, 'Tg C')
  }
  return(out)
}

terra_max = function(x) x@ptr$range_max
terra_min = function(x) x@ptr$range_min

calc_fill = function(x) {
  mx = max(x) %>% 
    ceiling
  mn = min(x) %>% 
    floor
  dif = mx - mn
  if (dif < 5) {
    out = seq(mn, mx, by=1)
  } else if (dif < 10) {
    out = seq(mn, mx, by=2)
  } else if (dif < 50) {
    out = seq(10*floor(mn/10), 10*ceiling(mx/10), by=10)
  } else if (dif < 80) {
    out = seq(10*floor(mn/10), 10*ceiling(mx/10), by=15)
  } else if (dif < 100) {
    out = seq(10*floor(mn/10), 10*ceiling(mx/10), by=20)
  } else if (dif < 500) {
    out = seq(100*floor(mn/100), 100*ceiling(mx/100), by=100)
  }
  return(out)
}

mx_stock = max(
  terra_max(pltdf$CSTOCK_80), 
  terra_max(pltdf$CSTOCK_NOW))
mn_stock = min(
  terra_min(pltdf$CSTOCK_80),
  terra_min(pltdf$CSTOCK_NOW)
)

fill_limits = rbind(
  CSTOCK_NOW = c(mn_stock, mx_stock),
  CSTOCK_80 = c(mn_stock, mx_stock),
  CGAP = c(0, ceiling(terra_max(pltdf$CGAP))),
  CSTOCK_NOW_se = c(0, ceiling(terra_max(data$CSTOCK_NOW_se))),
  CSTOCK_80_se = c(0, ceiling(terra_max(data$CSTOCK_NOW_se))),
  CGAP_se = c(0, ceiling(terra_max(data$CSTOCK_NOW_se)))
)
calc_fill = function(x) {
  mx = max(x)
  if (mx < 5) {
    out = seq(0, ceiling(mx), by=1)
  } else if (mx < 10) {
    out = seq(0, ceiling(mx), by=2)
  } else if (mx < 25) {
    out = seq(0, 10*ceiling(mx/10), by=5)
  } else if (mx < 50) {
    out = seq(0, 10*ceiling(mx/10), by=10)
  } else if (mx < 200) {
    out = seq(0, 10*ceiling(mx/10), by=30)
  } else if (mx < 500) {
    out = seq(0, 100*ceiling(mx/100), by=100)
  }
  return(out)
}

# this is a pretty slow function
subtitles = lapply(nn, function(x) generate_subtitle(data, x)) %>% 
  set_names(nn)
subtitles
```

## Plot
```{r fig.height=8}
fig4 = lapply(nn, function(x) {
  out = ggplot(mwi_utm) +
    # political boundaries as background
    scale_x_continuous(breaks=seq(33, 36, by=1)) +
    geom_sf(color='darkgray',
            fill='darkgray') +
    
    # stock level
    # new_scale('fill') +
    geom_stars(data=st_as_stars(pltdf[[x]])) +
    scale_fill_distiller(palette='YlGnBu',
                         direction=1,
                         na.value=NA,
                         limits=fill_limits[x,],
                         breaks=calc_fill(fill_limits[x,])) +
  
    # political boundaries
    geom_sf(color='white',
            fill=NA,
            size=0.3) +
    labs(title=prettynames[x],
         subtitle=subtitles[[x]],
         fill=expression(paste('Mg C ha'^{-1})))

  out = gg_mapthemer(out)+
    theme(legend.position=c(0.85, 0.7),
          axis.title=element_blank(),
          axis.text=element_text(size=8),
          legend.text=element_text(size=8),
          legend.title=element_text(size=8),
          plot.title=element_text(size=10),
          plot.subtitle=element_text(size=9),
          legend.key.width=unit(0.2, 'in'),
          legend.key.height=unit(0.15, 'in')
          )
  
  return(out)
})
```

```{r fig.height=5, fig.width=6.5}
grid.arrange(grobs=fig4, nrow=1)
```

## Export
```{r}
here(out_dir, 'Figure 4 - Maps of Stocks.jpg') %>% 
  jpeg(height=5, width=6.5, units='in', res=300)
grid.arrange(grobs=fig4, nrow=1)
dev.off()
```


# Figure S3 - Errors of Current Stocks
```{r fig.height=8}
pltdf_se = out_stocks[[is_se]] %>% 
  aggregate(fact=10) %>% 
  project(set_projection)

nn = names(pltdf_se)
```

## Plot
```{r}
fig4err = lapply(nn, function(x) {
  out = ggplot(mwi_utm) +
    # political boundaries as background
    scale_x_continuous(breaks=seq(33, 36, by=1)) +
    geom_sf(color='darkgray',
            fill='darkgray') +
    
    # stock level
    # new_scale('fill') +
    geom_stars(data=st_as_stars(pltdf_se[[x]])) +
    scale_fill_distiller(palette='Reds',
                         direction=1,
                         na.value=NA,
                         breaks=seq(0, 20, by=2)) +
  
    # political boundaries
    geom_sf(color='white',
            fill=NA,
            size=0.3) +
    labs(title=prettynames[x],
         fill=expression(paste("Mg C ha"^{-1}))) 

  out = gg_mapthemer(out)+
    theme(legend.position=c(0.85, 0.7),
          axis.title=element_blank(),
          axis.text=element_text(size=8),
          legend.text=element_text(size=8),
          legend.title=element_text(size=8),
          plot.title=element_text(size=10),
          plot.subtitle=element_text(size=9),
          legend.key.width=unit(0.2, 'in'),
          legend.key.height=unit(0.15, 'in')
          )
  
  return(out)
})
```

```{r fig.height=5, fig.width=6.5}
grid.arrange(grobs=fig4err, nrow=1)
```

## Export
```{r}
here(out_dir, 'Figure S3 - Errors of Maps of Stocks.jpg') %>% 
  jpeg(height=5, width=6.5, units='in', res=300)
grid.arrange(grobs=fig4err, nrow=1)
dev.off()
```

# Figure S4 - iSDA Stocks
## Predict
### Derive qr equation 
Using sampling points. Predicting unadjusted iSDA stock potentials
from iSDA soil properties (previously, used soil properties to predict adjusted
isda stock potentials)
```{r}
C_form = potential_stocks[['top_stat']]$formula %>% 
  as.character %>% 
  .[3] %>% 
  paste('ISDA_CSTOCK_NOW', ., sep='~') %>% 
  formula

quantiles = c(0.8)
qr_mod = rq(C_form, data=pt, tau=quantiles, ci=TRUE)
summary(qr_mod)
```
### Predict
```{r}
pred = 'ISDA_CSTOCK_80'
model = qr_mod

# predict
new_layers = predict(isda, 
                     model,
                     fun=rast_predict_se)

names(new_layers) = c(
  pred,
  paste0(pred, '_se')
)

# Set floor
tt = 
  max(new_layers[[pred]], 
      isda[['ISDA_CSTOCK_NOW']])
names(tt) = pred


terra::add(isda) = tt
terra::add(isda) = new_layers[[2]]
```

```{r}
isda$ISDA_CGAP = isda$ISDA_CSTOCK_80 - isda$ISDA_CSTOCK_NOW
```

## Map
### stocks and gaps layers
```{r fig.height=8}
is_stock = (grepl('CSTOCK', names(isda)) | grepl('CGAP', names(isda))) & !(grepl('_se', names(isda)))
isda_stocks = isda[[is_stock]]
```


### Plotting parameters
```{r fig.height=8}
isda_utm = aggregate(isda_stocks, fact=10) %>% 
  project(set_projection)
nn = names(isda_utm)

prettynames = c(
  ISDA_CSTOCK_NOW = 'iSDA Current Stock',# [Mg C/ha]',
  ISDA_CSTOCK_80 = 'iSDA Potential Stock',
  ISDA_CGAP = 'iSDA Current Gap'
)


generate_subtitle = function(data, x) {
  total = data[[x]] %>% 
    multiply_by(cellsize) %>% 
    global(fun='sum', na.rm=T) %>% 
    divide_by(1e6) %>% 
    round(1) 
    out = paste(total, 'Tg C')
  return(out)
}

#Skip setting fill limits to keep same as the adjsutment
isda_fill_limits = fill_limits
rownames(isda_fill_limits) %<>% paste('ISDA', ., sep='_')
# fill_limits = rbind(
#   ISDA_CSTOCK_NOW = c(mn_stock, mx_stock),
#   ISDA_CSTOCK_80 = c(mn_stock, mx_stock),
#   ISDA_CGAP = c(0, ceiling(terra_max(stocks_utm$CGAP)))
# )
isda_subtitles = lapply(nn, function(x) generate_subtitle(isda_stocks, x)) %>% 
  set_names(nn)
isda_subtitles
```

### Plot
```{r fig.height=8}
figS4 = lapply(nn, function(x) {
  out = ggplot(mwi_utm) +
    # political boundaries as background
    scale_x_continuous(breaks=seq(33, 36, by=1)) +
    geom_sf(color='darkgray',
            fill='darkgray') +
    
    # stock level
    # new_scale('fill') +
    geom_stars(data=st_as_stars(isda_utm[[x]])) +
    scale_fill_distiller(palette='YlGnBu',
                         direction=1,
                         na.value=NA,
                         limits=isda_fill_limits[x,],
                         breaks=calc_fill(isda_fill_limits[x,])) +
  
    # political boundaries
    geom_sf(color='white',
            fill=NA,
            size=0.3) +
    labs(title=prettynames[x],
         subtitle=isda_subtitles[[x]],
         fill=expression(paste('Mg C ha'^{-1})))

  out = gg_mapthemer(out)+
    theme(legend.position=c(0.85, 0.7),
          axis.title=element_blank(),
          axis.text=element_text(size=8),
          legend.text=element_text(size=8),
          legend.title=element_text(size=8),
          plot.title=element_text(size=10),
          plot.subtitle=element_text(size=9),
          legend.key.width=unit(0.2, 'in'),
          legend.key.height=unit(0.15, 'in')
          )
  
  return(out)
})

```

```{r fig.height=5, fig.width=6.5}
grid.arrange(grobs=figS4, nrow=1)
```

### Export
```{r}
here(out_dir, 'Figure S4 - iSDA Stocks.jpg') %>% 
  jpeg(height=5, width=6.5, units='in', res=300)
grid.arrange(grobs=figS4, nrow=1)
dev.off()
```

# Fig S5 - Differences with predicted
## Calculate
```{r}
diff = out_stocks$CSTOCK_NOW - isda_stocks$ISDA_CSTOCK_NOW
diff$CSTOCK_80 = out_stocks$CSTOCK_80 - isda_stocks$ISDA_CSTOCK_80
diff$CGAP = out_stocks$CGAP - isda_stocks$ISDA_CGAP
```

## Map

```{r}
pltdf = aggregate(diff, fact=10) %>% 
  project(set_projection)

nn = names(pltdf)

prettynames = c(
  CSTOCK_NOW = 'Adjustment in Current Stock',# [Mg C/ha]',
  CSTOCK_80 = 'Adjustment in Potential Stock',
  CGAP = 'Adjustment in Current Gap'
)

mx_stock = max(
  terra_max(pltdf$CSTOCK_80), 
  terra_max(pltdf$CSTOCK_NOW))
mn_stock = min(
  terra_min(pltdf$CSTOCK_80),
  terra_min(pltdf$CSTOCK_NOW)
)

generate_subtitle = function(data, x) {
  total = data[[x]] %>% 
    multiply_by(cellsize) %>% 
    global(fun='sum', na.rm=T) %>% 
    divide_by(1e6) %>% 
    round(1) 
  
    out = paste(total, 'Tg C')
  # }
  return(out)
}

diff_fill_limits = rbind(
  CSTOCK_NOW = c(mn_stock, mx_stock),
  CSTOCK_80 = c(mn_stock, mx_stock),
  CGAP = c(floor(terra_min(diff$CGAP)), ceiling(terra_max(diff$CGAP)))
)

diff_subtitles = lapply(nn, function(x) generate_subtitle(diff, x)) %>% 
  set_names(nn)
diff_subtitles
```


```{r}
figS5 = lapply(nn, function(x) {
  out = ggplot(mwi_utm) +
    # political boundaries as background
    scale_x_continuous(breaks=seq(33, 36, by=1)) +
    geom_sf(color='darkgray',
            fill='darkgray') +
    
    # stock level
    # new_scale('fill') +
    geom_stars(data=st_as_stars(pltdf[[x]])) +
    scale_fill_distiller(palette='Reds',
                         direction=1,
                         na.value=NA,
                         limits=diff_fill_limits[x,],
                         breaks=calc_fill(diff_fill_limits[x,])) +
  
    # political boundaries
    geom_sf(color='white',
            fill=NA,
            size=0.3) +
    labs(title=prettynames[x],
         subtitle=diff_subtitles[[x]],
         fill=expression(paste('Mg C ha'^{-1})))

  out = gg_mapthemer(out)+
    theme(legend.position=c(0.85, 0.7),
          axis.title=element_blank(),
          axis.text=element_text(size=8),
          legend.text=element_text(size=8),
          legend.title=element_text(size=8),
          plot.title=element_text(size=10),
          plot.subtitle=element_text(size=9),
          legend.key.width=unit(0.2, 'in'),
          legend.key.height=unit(0.15, 'in')
          )
  
  return(out)
})
```

```{r fig.height=5, fig.width=6.5}
grid.arrange(grobs=figS5, nrow=1)
```

```{r}
here(out_dir, "Figure 5 - Difference in Stocks.jpg") %>% 
  jpeg(width=6.5, height=5, units='in', res=300)
grid.arrange(grobs=figS5, nrow=1)
dev.off()
```

