---
title: "5. Carbon Gap Implications"
knit: (function(inputFile, encoding) {
  rmarkdown::render(inputFile, encoding=encoding, output_dir=here::here('Results'))})
output:
  html_document:
    toc: true
    toc_depth: 5
    toc_float: true
    df_print: paged
    code_folding: hide
author: PME
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Interpretations
1. Time to saturation under 4 parts per mille improvement
2. Sequestration potential of non-responsive sites

# Load

## Paths, Variables, Libraries

Variables
```{r}
set_datum = 'epsg:4326'  # WGS84
set_projection = 'epsg:32736'  # utm 36S.
# output
overwrite = FALSE
# overwrite = TRUE

plot_diagnostics = TRUE
```

Paths
```{r}
# locations
in_dir = 'Data'
in_pt = 'Master Points.gpkg'

dir_spatial = 'Spatial' # in in_dir
out_dir = 'Results'

in_isda = 'ISDA_AR_MWI.tif'
in_suit = 'LI_ETAL_2017_MW_AGRI_LAND_SUIT.tif'  # cropland suitability
in_mwi = "Malawi gadm36_MWI.gpkg"
in_stocks = 'corrected_SOC_stocks_mwi.tif'
# in_dem = "SRTM90_V4_MWI.tif"  # from Google Earth Engine via ree

```

Libraries
```{r}
libs = c(
  # Coding helpers
  'here', 
  # 'tidyr',
  'magrittr',
  
  # spatial stuff
  'sf',
  'terra',
  'stars',
  
  # Other modeling
  'glmnet',
  # 'quantreg',
  
  # plotting
  'ggplot2',
  'rasterVis'
  # 'gridExtra'
)
```

```{r message=FALSE}
for (i in libs) {
  if (!require(i, character.only=TRUE)) {
    renv::install(i)
    require(i, character.only=TRUE)
}}

resource = function() {
  ll = here('Scripts', 'V3 Functions') %>% 
  list.files(full.names=TRUE)
  for (i in ll) {
    source(i)
  }
}
resource()
```

Versions
```{r}
sapply(libs, function(x) as.character(packageVersion(x))) %>% 
  as.matrix %>% 
  set_colnames(R.version.string)
```

## More functions
```{r}
gg_mapthemer = function(x, 
                        xlab=NULL, 
                        ylab=NULL,
                        coord_equal=FALSE) {
  out = x +
    labs(x=xlab,
         y=ylab) +
    theme_minimal() +
    theme(axis.text.x=element_text(angle=90, 
                                   hjust=1, 
                                   vjust=0.5))
  if (coord_equal) {
    out = out + coord_equal()
  }
  return(out)
}

ggmap = function(raster, fill='value', label=NULL) {
  require(rasterVis)
  out = gplot(raster) +
    geom_tile(aes_string(fill=fill)) +
    scale_fill_distiller(palette='YlGnBu', 
                        direction=1, 
                        na.value=NA) +
    coord_equal() +
    labs(fill=label)
  out = gg_mapthemer(out)
  return(out)
}
```

## Load
```{r}
load(here(in_dir, 'Fig 2 - Field Estimates.Rdata'))
load(here(in_dir, 'Part 3 - iSDA Corrections.Rdata'))
```

### Corrected stocks
```{r}
stocks = here(out_dir, in_stocks) %>%
  rast
ggmap(stocks$CSTOCK_NOW) +
  labs(title='Current Stocks [Mg/ha]')
```

### iSDA
load and convert units
```{r fig.height=4, fig.width=8}
isda = here(in_dir, dir_spatial, in_isda) %>% 
  rast %>% 
  mask(stocks$CSTOCK_NOW)
ggmap(isda$ISDA_SOC) +
  labs(title='iSDA SOC [g/kg]')
```

# Headline Values
## Sampled Fields
Average gap increase from adjusting iSDA on sampled fields
```{r}
c_to_co2e = function(x) x*44.01/12.011
co2e_to_dollar = function(x) c_to_co2e(x)*85 
get_totals = function(x) {
  c(
    carbon = x,
    co2 = c_to_co2e(x),
    dollars = co2e_to_dollar(x)
  )
}

avg_gap_increase = with(pt, CGAP - ISDA_CGAP) %>% 
  mean(na.rm=T)
get_totals(avg_gap_increase)
```

## Total dollar value
of sequestration
```{r}
map_total = function(x) {
  {x * cellSize(x, unit='ha')} %>% 
    global(fun='sum', na.rm=TRUE)
}
total_gap = map_total(stocks$CGAP)

get_totals(total_gap)
```

# Amounts across time
## First year sequestration potential
```{r}
rate = 0.004
total_emissions = 19.2*1e6 # in Mg (reported in Mt)

calc_seq_years = function(current, potential, rate, years) {
  out =
    {current * (1+rate)^years} %>% 
    min(potential)  %>% 
    {. - current}
}


timeseq_1 = calc_seq_years(stocks$CSTOCK_NOW, stocks$CSTOCK_80, rate, 1)

ggmap(timeseq_1, label='Carbon\n[Mg/ha]') +
  labs(title='C sequestered in first year')
```

```{r}
total_timeseq_1 = map_total(timeseq_1)
total_timeseq_1 %>% 
  get_totals
```
units are Mg

```{r}
c_to_co2e(total_timeseq_1)/total_emissions
```

## 50% of fields reach saturation
```{r}
calc_time_saturation = function(now, goal, rate) {
  out = log(goal/now, base=1+rate)
}

time_sat = calc_time_saturation(stocks$CSTOCK_NOW,
                                stocks$CSTOCK_80,
                                rate)
med_years = global(time_sat[[1]],
                   fun=median,
                   na.rm=T) %>% 
  ceiling
ggmap(time_sat, label='Years') +
  labs(title='Time to Saturation',
       subtitle=paste('Median:', med_years, 'years'))
```

```{r}
timeseq_med = calc_seq_years(stocks$CSTOCK_NOW, stocks$CSTOCK_80, rate, unlist(med_years))

ggmap(timeseq_med, label='Carbon\n[Mg/ha]') +
  labs(title='C sequestered by 50% saturation')
```

How much carbon is this?
```{r}
seq_median_sat = map_total(timeseq_med) %>% 
  unlist %>% 
  get_totals %T>%
  print
```



# Non responsive sites sequestration potential
Burke et al identified a threshold of 9.4 g carbon/kg soil for fertilizer responsiveness.

First find sites below that threshold. This requires updating iSDA carbon 
concentration using the relationship found with the bottom-up adjustment.

Then calculate time to the threshold. 
```{r}
thresh = .0094  #g C/kg soil
rate = 0.004 # percent increase per year

# correct iSDA SOC
correct_SOC = predict(isda,
                      current_stocks$`bottom totalc`,
                      type='response')

# identify pixels sites that are above the threshold
is_nonresponsive = correct_SOC <= thresh
is_nonresponsive[!is_nonresponsive] = NA

# Mask responsive pixels and calculate time to responsiveness
time_to_responsive = mask(correct_SOC, is_nonresponsive) %>% 
  calc_time_saturation(thresh, rate)
```


```{r}
ggmap(time_to_responsive, label='Years') +
  labs(title='Time to Responsiveness',
       subtitle=paste0(rate*100, '% annual increase'))
```
## Median time to responsiveness
```{r}
tt = spatSample(time_to_responsive, 1000, as.points=FALSE, na.rm=TRUE)
lattice::densityplot(~lyr.1, data=tt, main=paste("Median =", median(tt$lyr.1)))
```

## C sequestration to responsiveness
```{r}
cseq_to_responsive =
  mask(stocks$CSTOCK_NOW, 
       is_nonresponsive) %>% 
  calc_seq_years(stocks$CSTOCK_80, rate, time_to_responsive)


ggmap(cseq_to_responsive, label='SOC [Mg/ha]')

```

```{r}
total_cseq_to_responsive = map_total(cseq_to_responsive) %>% 
  get_totals %>% 
  print
```
```{r}
total_cseq_to_responsive %>% 
  sapply('/', 1e6)
```
Units are: Pg C, Mt CO2e, million USD 